
\chapter{Dot Carmichael}

\begin{enumerate}
  \item Most programming languages require the use of brackets to
    enclose the index in a reference to an element of an array.
  \begin{enumerate}
    \item Identify a language the requires the use of parentheses
      to enclose the index in a reference to an element of an array.
    \item Why did the designers of the language choose parentheses
      rather than brackets?
    \end{enumerate}

  \begin{answer}

  \begin{enumerate}
    \item Ada uses parentheses to enclose the index.
    \item The programmers used parentheses instead of brackets to provide
          uniformity between arrays and functions, since they both use
          mapping.
    \end{enumerate}

    \end{answer}
    
  \item What is the relationship between a lexeme and a token?

  \begin{answer}

    A token is a category of one or more lexemes within a language.

    \end{answer}

  \item
  \begin{enumerate}
    \item What kind of symbols are found at the internal nodes of a
      parse tree?
    \item What kind of symbols are found at the leaves of a parse tree?
    \end{enumerate}

  \begin{answer}

  \begin{enumerate}
    \item Non-terminal symbols.
    \item Terminal symbols.
    \end{enumerate}

    \end{answer}


  \item One of the most significant contributions from the developers
    of ALGOL 60 also limited the success of that language. What was
    that contribution?

  \begin{answer}

    Backus-Naur Form syntax description

    \end{answer}

  \item What problem were the creators of Common LISP trying to solve?

  \begin{answer}

    The problem was a lack of portability between programs written in
    different dialects of LISP.

    \end{answer}

  \item What is an ambiguous context free grammar?

  \begin{answer}

    An ambiguous context free grammar can generate sentences with more than
    one leftmost and rightmost derivation.

    \end{answer}

  \item Contrast the complexity of algorithms that can parse strings
    that conform to the most general kinds of context free grammars
    and the complexity of the algorithms that can parse strings that
    conform to the grammars of programming languages?

  \begin{answer}

    Algorithms for parsing any unambiguous grammars are highly inefficient,
    with a complexity of O(n**3), whereas algorithms for programming languages
    are complexity O(n).

    \end{answer}

  \item Java represents characters with Unicode. It is the first
    widely used programming language with this feature. What is the
    significance of this feature?

  \begin{answer}

    Unicode allows programs to be written which can communicate with other
    countries whose languages cannot be represented by the basic english
    alphabet.

    \end{answer}

  \item How does the binary coded decimal type differ from the
    floating point type?

  \begin{answer}

    Floating point numbers are an approximation of the number using fractions
    and exponents, stored in binary which can make them even less accurate.
    Binary coded decimals precisely store decimal values at the cost of taking
    up more storage than floating point numbers.

    \end{answer}

  \item Identify a user-defined ordinal type in the Java programming
    language.

  \begin{answer}

    Enumeration

    \end{answer}

  \item Mathematicians and programmers might have different ideas
    about the precedence of Boolean operators. Explain.

  \begin{answer}

    When Boolean operators evaluate to 0 and 1, they can be used in equations
    with multiple relational operators to produce unexpected results.
    (1 < 2 <= 1) -> ((true/1) <= 1) -> true

    \end{answer}

  \item Programmers should use \verb+===+ rather than \verb+==+ to
    test the equality of the values of two expressions in JavaScript. Why?

  \begin{answer}

    With +==+, strings can be coerced into numbers to check for equality.
    With +===+, no coercion will take place, so strings and numbers will not
    resolve as equal.

    \end{answer}

  \item Describe a hazard of allowing short-circuited evaluation
    of expressions and side effects in expressions at the same time.

  \begin{answer}

    If there is a side effect in the right hand side of the expression and
    it short-circuits on the left, then the side effect will not occur.

    \end{answer}

  \item Briefly describe the three steps in the mark-sweep algorithm
    for garbage collection.

  \begin{answer}

    1- all cells in heap set indicators to 'garbage'
    2- all pointers in the program are traced, and anything which is pointed
       to is changed to 'not garbage'
    3- all cells still marked 'garbage' are reverted to available space

    \end{answer}

  \item What led Yukihiro Matsumoto to create the Ruby programming language?

  \begin{answer}

    Matsumoto was dissatisfied with Perl and Python; specifically, that they
    were not pure object-oriented languages because they had primitive types
    and functions.

    \end{answer}

  \item What did Microsoft aim to achieve with its development of the
    C\# language?

  \begin{answer}

    Microsoft aimed to provide a language for development of component-based
    software in the .NET framework.

    \end{answer}

  \end{enumerate}



\section{More questions for discussion and review.}

\begin{enumerate}
  \item The design of which machine influenced the design
    of the control statements in FORTRAN?

  \begin{answer}

    IBM 704 system

  \end{answer}

  \item How many different kinds of control statements
    must the designer of a programming language include
    in a language?

  \begin{answer}

    While it is possible to use only one (GOTO), the
    minimum number of control statements in a language
    which does not use GOTO is two. One for choosing
    between control flow paths, and one for logically
    controlled iterations.

  \end{answer}

  \item What is the one question that applies in the
    design of all statements that allow selection or
    iteration?

  \begin{answer}

    Should the control structure have multiple entries?

  \end{answer}

  \item What is an advantage of requiring that
    the \textbf{then} and \textbf{else} clauses of
    an \textbf{if} statement be compound statements?

  \begin{answer}

    Requiring compound statements helps increase the
    readability and writability for programmers when
    using nested selector statements, that otherwise
    can get very messy and complicated.

  \end{answer}

  \item How does the \textbf{switch} statement in C\#
    differ from the \textbf{switch} statement in Java?

  \begin{answer}

    In Java, the switch statement does not allow case
    expressions anywhere except the top level in the body
    of the switch.  C\#  allows the execution of more than
    one segment. In C\#, the control expressions as well as
    case statements can be strings.

  \end{answer}

  \item Distinguish between 2 statements in Ruby
    that correspond to Java's \textbf{switch} statement.

  \begin{answer}

    Case expressions are the Ruby equivalent of Java's \textbf{switch}
    statement. One is semantically similar to nested if
    statements with case - when - then. The other is with
    boolean expressions being evaluated one at a time from top
    to bottom. The value of this case expression is equivalent
    to the value of the \textbf{then} statement corresponding to the
    first \textit{true} \textbf{when} statement.

  \end{answer}

  \item Features of a programming language sometimes persist
    longer than a feature of computing hardware that inspired
    and supported that part of the language's design.
    Similarly, features of hardware sometimes persist longer
    than some parts of a language's design that were created
    to take advantage of that feature in hardware.

    Give examples.

  \begin{answer}

    \begin{enumerate}
      \item The IBM 704 influenced the design of control statements
      which are still used today, and prompted the development
      of Fortran.
      \item The Algol standard used several different syntaxes which,
      among other things, allowed Europeans to use a comma to
      denote a decimal point, while Americans could continue to
      use a period.
      \item Most of the popular languages of the past 50 years have
      been designed around the von Neumann architecture. These
      are called imperative languages. In a von Neumann computer,
      both data and programs are stored in the same memory. The
      CPU, which executes instructions, is separate from the
      memory. Therefore, instructions and data must be
      transmitted, or piped, from memory to the CPU. Results of
      operations in the CPU must be moved back to memory. Nearly
      all digital computers built since the 1940s have been based
      on the von Neumann architecture.
      \item The \textbf{register} keyword in C is a hint to the compiler that
      a variable will be used repeatedly, and so it should be
      stored in the CPU rather than in memory. However, modern
      compilers are far better at optimization than programmers,
      so this keyword is outdated and unnecessary.
    \end{enumerate}

  \end{answer}

  \item Who most famously warned of the dangers of using the
    \textbf{goto} statement? What did Donald Knuth have to
    say about the use of the \textbf{goto} statement?

  \begin{answer}

    Edsger Dijkstra noted “The goto statement as it stands is
    just too primitive; it is too much an invitation to make a
    mess of one’s program.” Donald Knuth argued there were
    occasions when the efficiency of the goto outweighed its
    harm to readability.

  \end{answer}

  \item Describes Ada's \textbf{for} loop. Are there some
    kinds of iteration that might be easier in Ada than
    in Java? Easier in Java than in Ada?

  \begin{answer}

    The for loop looks like this:
    \begin{lstlisting}
      for variable in [reverse] discrete_range loop … end loop;
    \end{lstlisting}
    Ada’s for loop can use any ordinal type variable for
    its counter. Arrays with ordinal type subscripts can be
    conveniently processed.

  \end{answer}

  \item What does it mean to say that the guarded commands
    of Ada are non-deterministic?

  \begin{answer}

    Guarded commands in Ada are nondeterministically chosen
    for execution when more than one of the statements are
    evaluated to true. This means that if there are three
    guarded statements and two of the three evaluate to true,
    then each time the program will use one of the two
    statements. It will not always use the one that appears
    first nor the one that appears last, but rather it will
    choose between them non-deterministically or randomly.

  \end{answer}

  \item The header files in a C program contain function
    prototypes. What is a function prototype?

  \begin{answer}

    A function prototype is a function declaration that gives
    the function's name and type signature, but does not
    specify the function body. It is also referred to as a
    function interface at times. In other languages these are
    uncommon because subprograms do not need declarations since
    they do no need to be defined before they are called.

  \end{answer}

  \item Every method in a Ruby program belongs to a class.
    A programmer can place a definition of a method inside
    the definition of a class or outside of the definition
    of any class that the programmer writes. To which class
    does the method belong in the second case?

  \begin{answer}

    If a method is defined outside of the definition of any
    class that the programmer writes then the method belongs
    to the root object, \textbf{Object}.

  \end{answer}

  \item Distinguish between positional and keyword parameters.

  \begin{answer}

    Positional parameters are bound based on the order in which
    they are given to the function. Keyword parameters are when
    the name of the formal parameter to which an actual
    parameter is to be bound is specified with the actual
    parameter in a call. This allows them to be input in any
    order.

  \end{answer}

  \item Ruby blocks are closures. What does that mean?

  \begin{answer}

    A closure is an anonymous function which can be passed as
    a parameter. Also, Ruby doesn't nest scope, so variables
    defined in a method are not accessible outside of the
    method (say, in the class which defines them).

  \end{answer}

  \item What is a pure function?

  \begin{answer}

    A \textbf{pure function} is a \textbf{function} where the return value
    is only determined by its input values, without
    observable side effects.

  \end{answer}

  \item Some languages give programmers means to define
    both functions and procedures. Java doe not. Is that
    a serious limitation?

  \begin{answer}

    It does not seem to be a serious limitation, it just means
    that performing procedures in Java is more verbose because
    it requires that the programmer instantiate the class with
    any needed methods.

  \end{answer}

  \item Declarations of formal parameters in an Ada procedure
    can include, in addition to the names and types of the
    parameters, reserved words that do not appear in declarations
    in Java programs. 
    What is the purpose of those reserved words?

  \begin{answer}

    Ada allows the programmer to specify in mode, out mode,
    and inout mode for each formal parameter.  This means that
    they can receive data from the actual parameter, they can
    send data to the actual parameter, or they can do both.

  \end{answer}

  \item The C language imposes a constraint upon programmers
    who want to pass a multidimensional array to a function.
    What is the constraint? How did the design of the Java
    programming language eliminate that constraint for 
    programmers who use that language?

  \begin{answer}

    In C, programmers are required to pass the length of an
    array to a method. Specifying the length of an array
    (like in C) is unnecessary because Java arrays store a
    pointer to an address containing their length.

  \end{answer}

  \item An activation record contains a return
    address, a dynamic link, parameters, and
    local variables.
  \begin{enumerate}
    \item To what does the return address point?
    \item To what does the dynamic link point?
  \end{enumerate}

  \begin{answer}

    \begin{enumerate}
      \item The return address usually consists of a pointer
      to the instruction following the call in the code
      segment of the calling program unit.
      \item The dynamic link points to the base of the
      activation record instance of the caller.
    \end{enumerate}

  \end{answer}

  \item The stack will contain multiple activation
    records for a single subprogram under what
    circumstances?

  \begin{answer}

    In a recursive program there can be multiple activation
    records (although they will be incomplete) for a single
    subprogram.

  \end{answer}

  \item How (or why?) does the LIFO protocol apply to
    calls to and returns from subprograms?

  \begin{answer}

    LIFO allows for subprograms to be nested within one
    another and for subprograms to be used as parameters
    within other subprograms.

  \end{answer}

  \item Which important development in computer architecture
    has changed the way that the stack is used in some
    systems for facilitating calls to and returns from
    subprograms?

  \begin{answer}

    RISC (reduced instruction set computing) machines
    have parameters passed in registers in their
    compilers because RISC machines have more registers
    than CISC (complex instruction set computing)
    machines. Chapter 10 assumes parameters are passed
    in the stack though, as they had been in CISC
    machines.

  \end{answer}

  \item A dynamic chain contains a history of what?

  \begin{answer}

    Dynamic chain represents the dynamic history of how
    execution got to its current position, which is
    always in the subprogram code whose activation
    record instance is on top of the stack.

  \end{answer}

  \item Which two numbers are needed to compute
    the address of a local variable in a subprogram?

  \begin{answer}

    To compute the address of a local variable you need
    the (chain_offset, local_offset) pair.

  \end{answer}

  \item How does a Ruby module differ from a class?

  \begin{answer}

    Modules are unlike classes in that they cannot be
    instantiated or subclassed and do not define
    variables. Methods that are defined in a module
    include the module’s name in their names.

  \end{answer}

  \item Memory for variables can be allocated on the heap
    and on the stack. In which place or places is memory
    allocated for objects in C++? in Java?

  \begin{answer}

    In C++, variables can be allocated to the heap either
    by making them “static” or by allocating memory with
    the keyword “new”. However, variables which are
    initialized during the execution of a function are
    allocated to the stack. Java behaves the same way.

  \end{answer}

  \item What problems were solved by the addition
    of genericity to Java?

  \begin{answer}

    Generics allow a type or method to operate on objects
    of various types while providing compile-time type
    safety.

  \end{answer}

  \item What is the purpose of the static chain?

  \begin{answer}

    The static chain is a path of pointers which go
    from each function to its parent. They allow
    child subprograms to use variables which are
    local to their parent, grandparent, or farther
    up, without needing to copy those variables to
    the call stack.

  \end{answer}

  \item What is a singleton?

  \begin{answer}

    A singleton is a class which provides a global
    access point to a single instance. This is useful
    for tasks which only need one point of access,
    like a file system.

  \end{answer}

  \item What are the two parts of the definition 
    of an abstract data type?

  \begin{answer}

    \begin{enumerate}
      \item A type definition which allows program units
      to declare variables of the type but hides the
      representation of objects of the type.
      \item A set of operations for manipulating objects
      of the type.
    \end{enumerate}

  \end{answer}
\end{enumerate}